---
title: "Writing Package Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing Package Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(EpiForsk)
```

This vignettes contains hints and recommendations for writing R functions within
a package. If you are not already comfortable with writing R functions, you can
read the chapters from [R for data science](https://r4ds.had.co.nz/functions.html)
and [Advanced R](https://adv-r.hadley.nz/functions.html). 

## Setting up
Your R function must be contained in a .R file under the R folder. As a rule, 
one R script contains only one function. If you have functions that are
closely related, such as a function returning an S3 object and associated S3 
methods, it can make sense to have them all in the same script. We recommend 
combining functions that share common documentation. Combining the documentation
of multiple functions is achieved using the roxygen2 tag  `@rdname`.

## Dependencies
When writing an R function, you will likely call functions from other packages.
You may be used to calling `library()` or `require()` to load the namespace of a
package and attach it to the search path. Never do this when writing package
functions, as dependencies are handles elsewhere. Specifically, the DESCRIPTION
file contains an imports field, listing all external packages needed for our
package to work. Packages listed here will be installed when EpiForsk is
installed. To add a package to the imports, use `usethis::use_package()`. Note
the packages added to the description file are not imported into the EpiForsk
package. The imports field just ensures the packages are present on the users
computer. To use functions from external packages, use the `::` operator.
Writing `pkg::name` will load package `pkg` if it isn't already loaded and
access the function `name`. It is possible to import functions from other
packages into the EpiForsk namespace using `@importFrom` for a single function
or `@import` for the full package. It must be added to the roxygen2
documentation in the `EpiForsk_package.R` file. Doing this makes the imported
functions part of the package, increasing the chance of conflicting function
names as well as increasing the package size. Therefore `::` is the default way
to use external functions. We recommend using `@importFrom` for operators with
special syntax such as `%>%` and for functions used repeatedly. Calling `::`
costs around 5 `r knitr::asis_output("\U03bc")`s, so repeatedly means when the
function could potentially be called millions of times. We recommend avoiding
`@import`. For more, see 
[dependencies](https://r-pkgs.org/dependencies-mindset-background.html).

## Loading package functions
To test your function during development, you can use `devtools::load_all()`
to make it available in the R session. This makes testing smoother than
waiting for the package to install every time.
